## Ames Housing Data Set
```{r, include = FALSE}
# enable caching for this document
knitr::opts_chunk$set(cache = TRUE)
library(magrittr)
library(ggplot2)
library(tibble)

root = rprojroot::find_root(rprojroot::is_git_root)
ap = adjust_path(paste0(getwd(), "/figure"))
```

Ames is a small city in Iowa, USA.

It describes the sale of individual residential properties from 2006 to 2010.

The data was collected by Dean De Cock as an more complex alternative to the often used Boston Housing dataset.

It contains 2930 observations and

- 23 categorical features,
- 23 ordered categorical features,
- 12 integer features,
- 20 continuous features,
- 1 functional feature describing power usage of houses over a day.


__Note:__ This is a slightly changed dataset from the original one.

## Ames Housing Data Set - Target

The goal is to predict the selling price based on these features

```{r, echo=FALSE, message = FALSE, fig.height=5.5}
data = read.csv(paste0(root, "/data/ames_housing_extended.csv"))

ggplot(data, aes(x = SalePrice)) + geom_histogram() + theme_minimal()

```
