---
title: "Applied Machine Learning"
subtitle: "In-class exercise 04-X"
output:
  prettydoc::html_pretty:
    theme: hpstr
    toc: yes
    toc_depth: 2
    css: ../_template/prettydoc-hpstr.css
    self_contained: yes
---

```{r, include=FALSE}
sys.source("setup.R", envir = knitr::knit_global())
```

# Goal

After this exercise, you should be able to perform feature selection using wrapper functions with `mlr3fselect`. 

# Wrapper Methods

In addition to filtering, wrapper methods are another variant of selecting features. While in filtering conditions for the feature values are set, in wrapper methods the learner is applied to different subsets of the feature set. As models need to be refitted, this method is computationally expensive.

For wrapper methods, we need the package `mlr3fselect`, at whose heart the following `R6` classes are:

- `FSelectInstanceSingleCrit`, `FSelectInstanceMultiCrit`: These two classes describe the feature selection problem and store the results.
- `FSelector`: This class is the base class for implementations of feature selection algorithms.

# Prerequisites

We load the most important packages and use a fixed seed for reproducibility. 

```{r}
library(mlr3verse)
# library(mlbench)
library(data.table)
library(mlr3fselect)
set.seed(7832)
```

In this exercise, we will use the `german_credit` data and the learner `classif.rpart`:

```{r}
task_gc = tsk("german_credit")
lrn_rpart = lrn("classif.rpart")
```

# 1 Basic Application

## 1.1 Create the Framework

Create an `FSelectInstanceSingleCrit` object using `fsi()`. The instance should use a 3-fold cross validation, classification accuracy as the `measure` and terminate after 20 evaluations. For simplification only consider the features `age`, `amount`, `credit_history` and `duration`.

<details>
<summary>**Hint 1:**</summary>

```{r, eval = FALSE}
task_gc$select(...)

instance = fsi(
  task = ...,
  learner = ...,
  resampling = ...,
  measure = ...,
  terminator = ...
)
```

</details>

```{r, eval = !show.solution, echo = FALSE, results='asis'}
cat("<!--")
```

### Solution:
<details>
<summary>**Click me:**</summary>

```{r, eval = show.solution}
task_gc$select(c("age", "amount", "credit_history", "duration"))

instance = fsi(
  task = task_gc,
  learner = lrn_rpart,
  resampling = rsmp("cv", folds = 3),
  measure = msr("classif.acc"),
  terminator = trm("evals", n_evals = 20)
)
```

```{r, eval = !show.solution, echo = FALSE, results='asis'}
cat("-->")
```


## 1.2 Start the Feature Selection

Start the feature selection step by selecting `sequential` using the `FSelector` class via `fs()` and pass the `FSelectInstanceSingleCrit` object to the `$optimize()` method of the initialized `FSelector` object.

<details>
<summary>**Hint 1:**</summary>

```{r, eval = FALSE}
fselector = fs(...)
```

</details>

<details>
<summary>**Hint 2:**</summary>

```{r, eval = FALSE}
fselector = fs(...)
fselector$optimize(...)
```

</details>

```{r, eval = !show.solution, echo = FALSE, results='asis'}
cat("<!--")
```

### Solution:
<details>
<summary>**Click me:**</summary>

```{r, eval = show.solution}
fselector = fs("sequential")
fselector$optimize(instance)
```

The two calls (`fsi()`and `fs()`) can also be executed by one sugar function (`fselect()`): 

```{r, eval = FALSE}
instance = fselect(
  fselector = fs("sequential"),
  task =  tsk_gc,
  learner = lrn_rpart,
  resampling = rsmp("cv", folds = 3),
  measure = msr("classif.acc")
)
```

```{r, eval = !show.solution, echo = FALSE, results='asis'}
cat("-->")
```

## 1.3 Evaluate 

View the four characteristics and the accuracy from the instance archive for each of the first two batches. 

<details>
<summary>**Hint 1:**</summary>

```{r, eval = FALSE}
instance$archive$data[...]
```

</details>

<details>
<summary>**Hint 2:**</summary>

```{r, eval = FALSE}
instance$archive$data[batch_nr == ..., ...]
```

</details>

```{r, eval = !show.solution, echo = FALSE, results='asis'}
cat("<!--")
```

### Solution:
<details>
<summary>**Click me:**</summary>

```{r, eval = show.solution}
instance$archive$data[batch_nr == 1, 1:5]
```

The highest accuracy results when the model relies on `credit_history`. So for the next batch, only combinations with `credit_history` are considered:

```{r, eval = show.solution}
instance$archive$data[batch_nr == 2, 1:5]
```

We see that the accuracy did not increase, so the prediction did not improve when using two features compared to `credit_history`. 

Alternatively with `data.table`:

```{r, eval = show.solution}
as.data.table(instance$archive)[batch_nr == 1, 1:5]
as.data.table(instance$archive)[batch_nr == 2, 1:5]
```

A visualization of this can be created by `autoplot`:

```{r, eval = show.solution}
autoplot(instance, type = "performance")
```

```{r, eval = !show.solution, echo = FALSE, results='asis'}
cat("-->")
```

## 1.4 Model Training

Which feature(s) should be selected? Train the model.

<details>
<summary>**Hint 1:**</summary>

Compare the accuracy values for the different feature combinations and select the feature(s) accordingly.

</details>

<details>
<summary>**Hint 2:**</summary>

```{r, eval = FALSE}
task_gc = ...
task_gc$select(...)
lrn_rpart$train(...)
```

</details>

```{r, eval = !show.solution, echo = FALSE, results='asis'}
cat("<!--")
```

### Solution:
<details>
<summary>**Click me:**</summary>

The feature(s) resulting in the highest accuracy should be selected. Here, considering more features than `credit_history` does not improve the accuracy, so we only select this single feature:

```{r, eval = show.solution}
task_gc = tsk("german_credit")
task_gc$select(instance$result_feature_set)
lrn_rpart$train(task_gc)
lrn_rpart$selected_features()
```

```{r, eval = !show.solution, echo = FALSE, results='asis'}
cat("-->")
```

# 2 Multiple Performance Measures

To optimize multiple performance metrics, the same steps must be followed as above except that multiple metrics are passed. Create an ´instance´ object as above considering the measures `classif.tpr` and `classif.tnr`. For the second step use `random search` and take a look at the results in a third step.

We again use the `german_credit` data:

```{r}
task_gc = tsk("german_credit")
```

<details>
<summary>**Hint 1:**</summary>

```{r, eval = FALSE}
instance = fsi(...)
```

```{r, eval = FALSE}
fselector = fs(...)
fselector$...(...)
```

```{r, eval = FALSE}
features = unlist(lapply(...))
cbind(features,...)
```

</details>

```{r, eval = !show.solution, echo = FALSE, results='asis'}
cat("<!--")
```

### Solution:
<details>
<summary>**Click me:**</summary>

```{r, eval = show.solution, results='hide'}
instance = fsi(
  task = task_gc,
  learner = lrn_rpart,
  resampling = rsmp("cv", folds = 3),
  measure = msrs(c("classif.tpr", "classif.tnr")),
  terminator = trm("evals", n_evals = 20)
)
```

```{r, eval = show.solution, results='hide'}
fselector = fs("random_search")
fselector$optimize(instance)
```

```{r, eval = show.solution}
features = unlist(lapply(instance$result$features, paste, collapse = ", "))
cbind(features,instance$result[,c("classif.tpr", "classif.tnr")])
```
Or with `data.table`:
```{r, eval = show.solution, results='hide'}
as.data.table(instance$result)[, .(features, classif.tpr, classif.tnr)]
```

Note that the measures can not be optimal at the same time so one has to choose according to their preferences. Here, we see different tradeoffs of sensitivity and specificity but no feature subset is dominated by another, i.e. has worse sensitivity and specificity than any other subset.

```{r, eval = !show.solution, echo = FALSE, results='asis'}
cat("-->")
```

# 3 Nested Resampling

Nested resampling enables finding unbiased performance estimators for the selection of features. In `mlr3` this is possible with the class `AutoFSelector`, whose instance can be created by the function `auto_fselector()`.

## 3.1 Create an `AutoFSelector` Instance

Implement an `AutoFSelector` object that uses random search to find a feature selection that gives the highest accuracy for a logistic regression with holdout resampling. It should terminate after 10 evaluations.

<details>
<summary>**Hint 1:**</summary>

```{r, eval = FALSE}
afs = auto_fselector(
  fselector = ...,
  learner = ...,
  resampling = ...,
  measure = ...,
  terminator = ...
)
```

</details>

```{r, eval = !show.solution, echo = FALSE, results='asis'}
cat("<!--")
```

### Solution:
<details>
<summary>**Click me:**</summary>

```{r, eval = show.solution}
afs = auto_fselector(
  fselector = fs("random_search"),
  learner = lrn("classif.log_reg"),
  resampling = rsmp("holdout"),
  measure = msr("classif.acc"),
  terminator = trm("evals", n_evals = 10)
)
afs
```

```{r, eval = !show.solution, echo = FALSE, results='asis'}
cat("-->")
```

## 3.2 Benchmark

Compare the `AutoFSelector` with a normal logistic regression using 3 fold cross-validation.

<details>
<summary>**Hint 1:**</summary>

The `AutoFSelector` inherits from the `Learner` base class, which is why it can be used like any other learner.

</details>

<details>
<summary>**Hint 2:**</summary>

Implement a benchmark grid and aggregate the result.

</details>

```{r, eval = !show.solution, echo = FALSE, results='asis'}
cat("<!--")
```

### Solution:
<details>
<summary>**Click me:**</summary>

```{r, eval = show.solution, results='hide'}
grid = benchmark_grid(tsk("sonar"), list(afs, lrn("classif.log_reg")),
  rsmp("cv", folds = 3))

bmr = benchmark(grid)$aggregate(msr("classif.acc"))
```

```{r, eval = show.solution}
as.data.table(bmr)[, .(learner_id, classif.acc)]
```

```{r, eval = !show.solution, echo = FALSE, results='asis'}
cat("-->")
```

# Summary

- Stratified resampling helps with balancing classes and features within CV folds, to ensure each fold represents the data well enough.
- Block resampling reduces bias in generalization error estimates by ensuring that observations from the same group end up in the same fold.
- Custom domain applications require custom performance measures. In `mlr3`, you can define custom measures by creating a new `R6` class.
