---
title: "Applied Machine Learning"
subtitle: "In-class exercise 10-2"
output:
  prettydoc::html_pretty:
    theme: hpstr
    toc: yes
    toc_depth: 2
    css: ../_template/prettydoc-hpstr.css
    self_contained: yes
---

```{r, include=FALSE}
sys.source("setup.R", envir = knitr::knit_global())
```

# Goal

Learn how to do stacking with `mlr3pipelines`.

# German Credit Data

## Description

- Data from 1973 to 1975 from a large regional bank in southern Germany classifying credits described by a set of attributes to good or bad credit risks.
- Stratified sample of 1000 credits (300 bad ones and 700 good ones).
- Customers with good credit risks perfectly complied with the conditions of the contract while customers with bad credit risks did not comply with the contract as required.
- Available in `tsk("german_credit")`.

## Data Dictionary

n = 1,000 observations of credits

- `credit_risk`: Has the credit contract been complied with (good) or not (bad)?
- `age`: Age of debtor in years
- `amount`: Credit amount in DM
- `credit_history`: History of compliance with previous or concurrent credit contracts
- `duration`: Credit duration in months
- `employment_duration`: Duration of debtor's employment with current employer
- `foreign_worker`: Whether the debtor is a foreign worker
- `housing`: Type of housing the debtor lives in
- `installment_rate`: Credit installments as a percentage of debtor's disposable income
- `job`: Quality of debtor's job
- `number_credits`: Number of credits including the current one the debtor has (or had) at this bank
- `other_debtors`: Whether there is another debtor or a guarantor for the credit
- `other_installment_plans`: Installment plans from providers other than the credit-giving bank
- `people_liable`: Number of persons who financially depend on the debtor
- `personal_status_sex`: Combined information on sex and marital status
- `present_residence`: Length of time (in years) the debtor lives in the present residence
- `property`: The debtor's most valuable property
- `purpose`: Purpose for which the credit is needed
- `savings`: Debtor's saving
- `status`: Status of the debtor's checking account with the bank
- `telephone`: Whether there is a telephone landline registered on the debtor's name

# Exercises

## Exercise 1: Single- and Multilayer Stacking

We now want to do stacking.
Use six learners on level 0: A decision tree, k-NN (with k = 7), an elastic net (alpha = 0.5 and s = 0.1) with categorical features target encoded, naive bayes with categorical features target encoded, XGBoost (nrounds = 100) with categorical features target encoded and a random forest (num.trees = 100).
Do not cross-validate on level 0 for now.
Train on the task and inspect the output of the level 0 graph.
Now combine level 0 with a logistic regression on level 1.
Proceed to benchmark the stacked ensemble against each level 0 learner and the random forest.
Use 5-fold CV to evaluate the ROC AUC of the models.

So far, we have not cross-validated the level 0 predictions.
Create another stacked ensemble where you use 3-fold CV to cross-validate the level 0 predictions.

So far, we have dropped the original features of the task.
Create another stacked ensemble where you additionally keep the original features of the task.

Finally, create a multi-layer stacking ensemble by passing the output of the current level 0 into a level 1 that is built similar as level 1.
Then use a random forest on the final level 2.
Make sure that when you pass the original features of the task to the final level 2 that you do not also pass the cross-validated predictions from level 1.

<details>
  <summary>**Hint 1:**</summary>
  To wrap a learner within an ensemble and use its predictions, the learner_cv `PipeOp` is essential.
  Its `resampling.method` hyperparameter allows you to cross-validate the predictions or not.
  Similarly `resampling.folds` allows you to specify the number of folds if predictions should be cross-validated.
  To combine the level 0 output, use the featureunion `PipeOp`.
  To keep the original features of the task in the output of level 0, use the nop `PipeOp` along the learners.
  To make sure that when you pass the original features of the task to the final level 2 that you do not also pass the cross-validated predictions from level 1, use an appropriately configured select `PipeOp`.
  The following gallery post could be helpful: https://mlr-org.com/gallery/pipelines/2020-04-27-tuning-stacking/index.html
</details>

<details>
  <summary>**Hint 2:**</summary>
 
```{r, eval = FALSE}
library(mlr3)
library(mlr3learners)
library(mlr3pipelines)
library(mlr3viz)

task = tsk("german_credit")

set.seed(2409)
resampling = rsmp("cv", folds = ...)$instantiate(task)
dt = lrn("classif.rpart")
dt$predict_type = "prob"
kknn = lrn(..., k = ...)
kknn$predict_type = "prob"
elnet = lrn(..., alpha = ..., s = ...)
elnet$predict_type = "prob"
elnet = as_learner(po(...) %>>% ...)
naive_bayes = lrn(...)
naive_bayes$predict_type = "prob"
naive_bayes = as_learner(po(...) %>>% ...)
xgboost = lrn(..., nrounds = ...)
xgboost = as_learner(po(...) %>>% ...)
xgboost$predict_type = "prob"
rf = lrn(..., num.trees = ..._
rf$predict_type = "prob"

level0 = gunion(
  list(
    po("learner_cv", learner = ..., resampling.method = ...),
    ...
  )
) %>>% po(...)
level0$plot()

level0_train_output = level0$train(task)
level0_train_output

level1 = lrn(..., id = "log_reg_out")
level1$predict_type = "prob"

ensemble = as_learner(level0 %>>% level1)
ensemble$id = "Simple Stacked Ensemble Insample"
ensemble$graph$plot()

bg = benchmark_grid(..., list(...), ...)
b = benchmark(bg)
autoplot(..., measure = ...

.
.
.
```
</details>

```{r, eval = !show.solution, echo = FALSE, results='asis'}
cat("<!--")
```
### Solution:

<details>
<summary>Click me:</summary>

```{r, eval = show.solution}
library(mlr3)
library(mlr3learners)
library(mlr3pipelines)
library(mlr3viz)

task = tsk("german_credit")

set.seed(2409)
resampling = rsmp("cv", folds = 5L)$instantiate(task)
dt = lrn("classif.rpart")
dt$predict_type = "prob"
kknn = lrn("classif.kknn", k = 7L)
kknn$predict_type = "prob"
elnet = lrn("classif.glmnet", alpha = 0.5, s = 0.01)
elnet$predict_type = "prob"
elnet = as_learner(po("encodeimpact") %>>% elnet)
naive_bayes = lrn("classif.naive_bayes")
naive_bayes$predict_type = "prob"
naive_bayes = as_learner(po("encodeimpact") %>>% naive_bayes)
xgboost = lrn("classif.xgboost", nrounds = 100L)
xgboost = as_learner(po("encodeimpact") %>>% xgboost)
xgboost$predict_type = "prob"
rf = lrn("classif.ranger", num.trees = 100L)
rf$predict_type = "prob"

level0 = gunion(
  list(
    po("learner_cv", learner = dt, resampling.method = "insample"),
    po("learner_cv", learner = kknn, resampling.method = "insample"),
    po("learner_cv", learner = elnet, resampling.method = "insample"),
    po("learner_cv", learner = naive_bayes, resampling.method = "insample"),
    po("learner_cv", learner = xgboost, resampling.method = "insample"),
    po("learner_cv", learner = rf, resampling.method = "insample")
  )
) %>>% po("featureunion")
level0$plot()

level0_train_output = level0$train(task)
level0_train_output
# This is a task that contains the probability predictions for each class from each learner
# To also have the predicted class, we could set the `resampling.keep_response` of the learner_cv `PipeOp`

level1 = lrn("classif.log_reg", id = "log_reg_out")
level1$predict_type = "prob"

ensemble = as_learner(level0 %>>% level1)
ensemble$id = "Simple Stacked Ensemble Insample"
ensemble$graph$plot()

bg = benchmark_grid(task, list(ensemble, dt, kknn, elnet, naive_bayes, xgboost, rf), resampling)
b = benchmark(bg)
autoplot(b, measure = msr("classif.auc"))

level0 = gunion(
  list(
    po("learner_cv", learner = dt, resampling.method = "cv", resampling.folds = 3L),
    po("learner_cv", learner = kknn, resampling.method = "cv", resampling.folds = 3L),
    po("learner_cv", learner = elnet, resampling.method = "cv", resampling.folds = 3L),
    po("learner_cv", learner = naive_bayes, resampling.method = "cv", resampling.folds = 3L),
    po("learner_cv", learner = xgboost, resampling.method = "cv", resampling.folds = 3L),
    po("learner_cv", learner = rf, resampling.method = "cv", resampling.folds = 3L)
  )
) %>>% po("featureunion")

ensemble_cv = as_learner(level0 %>>% level1)
ensemble_cv$id = "Simple Stacked Ensemble 3-fold CV"

bg = benchmark_grid(task, list(ensemble, ensemble_cv, rf), resampling)
b = benchmark(bg)
autoplot(b, measure = msr("classif.auc"))

level0 = gunion(
  list(
    po("learner_cv", learner = dt, resampling.method = "cv", resampling.folds = 3L),
    po("learner_cv", learner = kknn, resampling.method = "cv", resampling.folds = 3L),
    po("learner_cv", learner = elnet, resampling.method = "cv", resampling.folds = 3L),
    po("learner_cv", learner = naive_bayes, resampling.method = "cv", resampling.folds = 3L),
    po("learner_cv", learner = xgboost, resampling.method = "cv", resampling.folds = 3L),
    po("learner_cv", learner = rf, resampling.method = "cv", resampling.folds = 3L),
    po("nop")
  )
) %>>% po("featureunion")
level0_train_output = level0$train(task)[[1L]]
level0_train_output
# The task now also contains the original features of the task passed through

ensemble_cv_pass = as_learner(level0 %>>% level1)
ensemble_cv_pass$id = "Simple Stacked Ensemble 3-fold CV + Orig Features"

bg = benchmark_grid(task, list(ensemble, ensemble_cv, ensemble_cv_pass, rf), resampling)
b = benchmark(bg)
autoplot(b, measure = msr("classif.auc"))
# Usefulness of passing through the original features of the task usually strongly depends on the higher level learners
# If you have time, you may want to play around with the choice of the level 1 learner (or optimize it directly)

level0 = gunion(
  list(
    po("learner_cv", id = "dt1_0", learner = dt, resampling.method = "cv", resampling.folds = 3L),
    po("learner_cv", id = "kknn_0", learner = kknn, resampling.method = "cv", resampling.folds = 3L),
    po("learner_cv", id = "elnet_0", learner = elnet, resampling.method = "cv", resampling.folds = 3L),
    po("learner_cv", id = "naive_bayes_0", learner = naive_bayes, resampling.method = "cv", resampling.folds = 3L),
    po("learner_cv", id = "xgboost_0", learner = xgboost, resampling.method = "cv", resampling.folds = 3L),
    po("learner_cv", id = "rf_0", learner = rf, resampling.method = "cv", resampling.folds = 3L),
    po("nop", id = "nop_0")
  )
) %>>% po("featureunion", id = "featureunion_0")
level1 = gunion(
  list(
    po("learner_cv", id = "dt1_1", learner = dt, resampling.method = "cv", resampling.folds = 3L),
    po("learner_cv", id = "kknn_1", learner = kknn, resampling.method = "cv", resampling.folds = 3L),
    po("learner_cv", id = "elnet_1", learner = elnet, resampling.method = "cv", resampling.folds = 3L),
    po("learner_cv", id = "naive_bayes_1", learner = naive_bayes, resampling.method = "cv", resampling.folds = 3L),
    po("learner_cv", id = "xgboost_1", learner = xgboost, resampling.method = "cv", resampling.folds = 3L),
    po("learner_cv", id = "rf_1", learner = rf, resampling.method = "cv", resampling.folds = 3L),
    po("nop", id = "nop_1") %>>% po("select", selector = selector_name(task$feature_names))
  )
) %>>% po("featureunion", id = "featureunion_1")
level2 = lrn("classif.log_reg", id = "log_reg_out")
level2$predict_type = "prob"

ensemble_multilayer_cv_pass = as_learner(level0 %>>% level1 %>>% level2)
ensemble_multilayer_cv_pass$id = "Multilayer Stacked Ensemble 3-fold CV + Orig Features"
ensemble_multilayer_cv_pass$graph$plot()

bg = benchmark_grid(task, list(ensemble, ensemble_cv, ensemble_cv_pass, ensemble_multilayer_cv_pass, rf), resampling)
b = benchmark(bg)
autoplot(b, measure = msr("classif.auc"))
# Sometimes beating a random forest baseline can be really difficult
```
```{r, eval = !show.solution, echo = FALSE, results='asis'}
cat("-->")
```

## Exercise 2: Tuning a Stacked Ensemble

We now want to tune one of our stacked ensembles from above.
Use six learners on level 0: A decision tree, k-NN, an elastic net (alpha = 0.5) with categorical features target encoded, naive bayes with categorical features target encoded, xgboost (nrounds = 100) with categorical features target encoded and a random forest.
Use 3-fold CV to cross-validate the level 0 predictions and pass them into a logistic regression on level 1.
Tune the following hyperparameters of the following learners on level 0:

* Decision tree: `cp` from 1e-3 to 0.1 on log scale
* k-NN: `k` from 1 to 10
* Elastic net: `s` from 1e-5 to 1000 on log scale
* XGBoost: `eta` from 1e-5 to 1 on log scale
* random forest: `mtry.ratio` from 0.1 to 1

Construct a `TuningInstance` that terminates after 20 evaluations and inspect the search space.
First, use the same resampling as used in the previous exercise and optimize ROC AUC via Bayesian Optimization.
Use an initial design size of 10.
Optimize the instance and inspect the results.

Finally, wrap everything in an `AutoTuner`.
Use 3-fold CV as a resampling within the `AutoTuner` and optimize ROC AUC.

Benchmark this automatically tuned stacked ensemble against the untuned stacked ensemble and the random forest from above.
This can take a while.
Proper parallelization can be helpful.

<details>
  <summary>**Hint 1:**</summary>
  Defining a search space automatically is possible by setting hyperparameters to a tune token via `to_tune`.
  If you are not familiar with BO, the following might be helpful: https://mlr3book.mlr-org.com/chapters/chapter5/advanced_tuning_methods_and_black_box_optimization.html#sec-bayesian-optimization
  Regarding parallelization, you may want to look at: https://mlr3book.mlr-org.com/chapters/chapter10/advanced_technical_aspects_of_mlr3.html#sec-nested-resampling-parallelization
</details>

<details>
  <summary>**Hint 2:**</summary>
 
```{r, eval = FALSE}
library(mlr3tuning)
library(mlr3mbo)

dt = lrn("classif.rpart", cp = ...)
dt$predict_type = "prob"
kknn = lrn(..., k = ...)
kknn$predict_type = "prob"
elnet = lrn(..., alpha = ..., s = ...)
elnet$predict_type = "prob"
elnet = as_learner(po(...) %>>% ...)
naive_bayes = lrn(...)
naive_bayes$predict_type = "prob"
naive_bayes = as_learner(po(...) %>>% ...)
xgboost = lrn(..., nrounds = ..., eta = ...)
xgboost = as_learner(po(...) %>>% ...)
xgboost$predict_type = "prob"
rf = lrn(..., num.trees = ..., mtry.ratio = ...)
rf$predict_type = "prob"

level0 = gunion(
  list(
    ...
  )
) %>>% po(...)
ensemble_tuned_cv = as_learner(level0 %>>% lrn(..., id = "log_reg_out", predict_type = "prob"))
ensemble_tuned_cv$id = "Simple Tuned Stacked Ensemble 3-fold CV + Orig Features"

instance = ti(
  ...
)
instance$search_space

tuner = tnr("mbo")
tuner$args = list(...)
tuner$optimize(...)

instance$archive$data

at = AutoTuner$new(
  ...
)

bg = benchmark_grid(..., list(...), ...)
b = benchmark(...)
autoplot(..., measure = ...)
```
</details>

```{r, eval = !show.solution, echo = FALSE, results='asis'}
cat("<!--")
```
### Solution:

<details>
<summary>Click me:</summary>

```{r, eval = show.solution}
library(mlr3tuning)
library(mlr3mbo)

dt = lrn("classif.rpart", cp = to_tune(p_dbl(lower = 1e-3, upper = 1, logscale = TRUE)))
dt$predict_type = "prob"
kknn = lrn("classif.kknn", k = to_tune(p_int(lower = 1L, upper = 10L)))
kknn$predict_type = "prob"
elnet = lrn("classif.glmnet", alpha = 0.5, s = to_tune(p_dbl(lower = 1e-5, upper = 1000, logscale = TRUE)))
elnet$predict_type = "prob"
elnet = as_learner(po("encodeimpact") %>>% elnet)
naive_bayes = lrn("classif.naive_bayes")
naive_bayes$predict_type = "prob"
naive_bayes = as_learner(po("encodeimpact") %>>% naive_bayes)
xgboost = lrn("classif.xgboost", nrounds = 100L, eta = to_tune(p_dbl(lower = 1e-5, upper = 1, logscale = TRUE)))
xgboost = as_learner(po("encodeimpact") %>>% xgboost)
xgboost$predict_type = "prob"
rf = lrn("classif.ranger", num.trees = 100L, mtry.ratio = to_tune(p_dbl(lower = 0.1, upper = 1)))
rf$predict_type = "prob"

level0 = gunion(
  list(
    po("learner_cv", learner = dt, resampling.method = "cv", resampling.folds = 3L),
    po("learner_cv", learner = kknn, resampling.method = "cv", resampling.folds = 3L),
    po("learner_cv", learner = elnet, resampling.method = "cv", resampling.folds = 3L),
    po("learner_cv", learner = naive_bayes, resampling.method = "cv", resampling.folds = 3L),
    po("learner_cv", learner = xgboost, resampling.method = "cv", resampling.folds = 3L),
    po("learner_cv", learner = rf, resampling.method = "cv", resampling.folds = 3L)
  )
) %>>% po("featureunion")
ensemble_tuned_cv = as_learner(level0 %>>% lrn("classif.log_reg", id = "log_reg_out", predict_type = "prob"))
ensemble_tuned_cv$id = "Simple Tuned Stacked Ensemble 3-fold CV + Orig Features"

instance = ti(
  task = task,
  learner = ensemble_tuned_cv,
  resampling = resampling,
  measures = msr("classif.auc"),
  terminator = trm("evals", n_evals = 20L)
)
instance$search_space

tuner = tnr("mbo")
tuner$args = list(init_design_size = 10L)
tuner$optimize(instance)

at = AutoTuner$new(
  tuner = tuner,
  learner = ensemble_tuned_cv,
  resampling = rsmp("cv", folds = 3L),
  measure = msr("classif.auc"),
  terminator = trm("evals", n_evals = 20L)
)

bg = benchmark_grid(task, list(ensemble_cv, at, lrn("classif.ranger", predict_type = "prob")), resampling)
b = benchmark(bg)
autoplot(b, measure = msr("classif.auc"))
# Sometimes beating a random forest baseline can be really, really difficult
# Try to play around with the structure of the stacked ensemble and see if you can beat the random forest
# Maybe try different learners, maybe use weighted model averaging in level 1 instead of the random forest (as in previous exercises)
# You could also try different preprocessing steps in front of each learner, e.g., dimensionality reduction or feature engineering
```
```{r, eval = !show.solution, echo = FALSE, results='asis'}
cat("-->")
```
# Summary

We have learnt how to build simple and more complex stacked ensembles and how to cross-validate predictions and pass the original features of the task.
We have also seen how to optimize the hyperparameters of such an ensemble and how to wrap the tuned and stacked ensemble in an `AutoTuner` which allows for fair benchmark.

