---
title: "Applied Machine Learning"
subtitle: "In-class exercise 10-3"
output:
  prettydoc::html_pretty:
    theme: hpstr
    toc: yes
    toc_depth: 2
    css: ../_template/prettydoc-hpstr.css
    self_contained: yes
---

```{r, include=FALSE}
sys.source("setup.R", envir = knitr::knit_global())
```

# Goal

Implement greedy ensemble selection from scratch.
Compare weighted model averaging (soft voting) to weighted majority voting (hard voting).

# German Credit Data

## Description

- Data from 1973 to 1975 from a large regional bank in southern Germany classifying credits described by a set of attributes to good or bad credit risks.
- Stratified sample of 1000 credits (300 bad ones and 700 good ones).
- Customers with good credit risks perfectly complied with the conditions of the contract while customers with bad credit risks did not comply with the contract as required.
- Available in `tsk("german_credit")`.

## Data Dictionary

n = 1,000 observations of credits

- `credit_risk`: Has the credit contract been complied with (good) or not (bad)?
- `age`: Age of debtor in years
- `amount`: Credit amount in DM
- `credit_history`: History of compliance with previous or concurrent credit contracts
- `duration`: Credit duration in months
- `employment_duration`: Duration of debtor's employment with current employer
- `foreign_worker`: Whether the debtor is a foreign worker
- `housing`: Type of housing the debtor lives in
- `installment_rate`: Credit installments as a percentage of debtor's disposable income
- `job`: Quality of debtor's job
- `number_credits`: Number of credits including the current one the debtor has (or had) at this bank
- `other_debtors`: Whether there is another debtor or a guarantor for the credit
- `other_installment_plans`: Installment plans from providers other than the credit-giving bank
- `people_liable`: Number of persons who financially depend on the debtor
- `personal_status_sex`: Combined information on sex and marital status
- `present_residence`: Length of time (in years) the debtor lives in the present residence
- `property`: The debtor's most valuable property
- `purpose`: Purpose for which the credit is needed
- `savings`: Debtor's saving
- `status`: Status of the debtor's checking account with the bank
- `telephone`: Whether there is a telephone landline registered on the debtor's name

# Exercises

## Exercise 1: Implement Greedy Ensemble Selection

In previous exercises we have seen how to do weighted model averaging and how to optimize weights numerically.
We now want to implement a popular method for optimizing the weights when performing model averaging: Greedy ensemble selection (GES)

First, read the original paper (https://www.cs.cornell.edu/~caruana/ctp/ct.papers/caruana.icml04.icdm06long.pdf) and make yourself familiar with the slides covering greedy ensemble selection.
Then, revisit the first exercise set.

We wanted to do model averaging of a decision tree, a k-NN (k = 7) and a logistic regression.
To optimize the weights numerically, we wrote an objective function, that allows for optimizing the weights of each model so that the resulting weighted model averaging is optimal.

We now want to use GES instead of the CMA-ES.
First, we will implement GES without relying on the `bbotk` infrastructure.
Afterwards, you are free to take a look at how `Optimizer`s are implemented in `bbotk`, see, e.g., https://github.com/mlr-org/bbotk/blob/70f3f96ff0023066aa8eaa7c97484f0d6e518663/R/OptimizerCmaes.R
and implement greedy ensemble selection as a proper `Optimizer`.

In this exercise, we will also use an outer test split to properly evaluate the performance of our ensemble with optimized weight in an unbiased manner.

<details>
  <summary>**Hint 1:**</summary>
  GES is conceptually very simple:
  - You start with the empty ensemble and add the model that has highest validation performance.
  - Then, you check for each model what the validation performance of the ensemble would be if you add that model to the ensemble.
  - To calculate the weights for each model, you simply count the number of the times the model has been added to the ensemble so far and divide it by the total number of models added.
  - Finally, to calculate the probability predictions of the ensemble, you perform weighted model averaging as usual.
  - You repeat this until a given number of iterations has been reached.

</details>

<details>
  <summary>**Hint 2:**</summary>
 
```{r, eval = FALSE}
library(mlr3)
library(mlr3learners)
library(mlr3pipelines)
library(mlr3viz)
library(bbotk)
library(mlr3misc)

task = tsk("german_credit")

outer_split = rsmp("holdout")$instantiate(task)

### old code part from previous exercise using CMA-ES to numerically optimize the weights
### modified to work with the newly introduced outer split

set.seed(2409)
dt = lrn("classif.rpart")
dt$predict_type = "prob"
kknn = lrn("classif.kknn", k = 7L)
kknn$predict_type = "prob"
log_reg = lrn("classif.log_reg")
log_reg$predict_type = "prob"
featureless = lrn("classif.featureless")
featureless$predict_type = "prob"

gr = gunion(list(dt, kknn, log_reg)) %>>% po("classifavg")
grl = as_learner(gr)

bg = benchmark_grid(task, list(grl, dt, kknn, log_reg, featureless), outer_split)
b = benchmark(bg)
b$aggregate(msr("classif.auc"))

# optimize weights on the train/validation split from the outer_split
train_valid_task = task$clone(deep = TRUE)$filter(outer_split$train_set(1L))
k = 10L
resampling = rsmp("cv", folds = k)$instantiate(train_valid_task)

dt_pred = resample(train_valid_task, dt, resampling)
kknn_pred = resample(train_valid_task, kknn, resampling)
log_reg_pred = resample(train_valid_task, log_reg, resampling)

objective_function = function(xs) {
  weights = unlist(xs)
  weights = weights / sum(weights)  # sum to 1 normalization
  aucs = map_dbl(seq_len(k), function(fold) {
    dt_p = dt_pred$predictions()[[fold]]
    kknn_p = kknn_pred$predictions()[[fold]]
    log_reg_p = log_reg_pred$predictions()[[fold]]
    row_ids = dt_p$row_ids
    stopifnot(all(row_ids == kknn_p$row_ids) && all(row_ids == log_reg_p$row_ids))
    truth = dt_p$truth
    weighted_probs = Reduce("+", list(dt_p$prob * weights[1L], kknn_p$prob * weights[2L], log_reg_p$prob * weights[3L]))
    weighted_response = colnames(weighted_probs)[apply(weighted_probs, MARGIN = 1L, FUN = which.max)]
    weighted_p = PredictionClassif$new(row_ids = row_ids, truth = truth, response = weighted_response, prob = weighted_probs)
    weighted_p$score(msr("classif.auc"))
  })
  list(classif.auc = mean(aucs), weights = list(weights))
}

domain = ps(w_dt = p_dbl(0, 1), w_kknn = p_dbl(0, 1), w_log_reg = p_dbl(0, 1))
codomain = ps(classif.auc = p_dbl(0.5, 1, tags = "maximize"))

objective = ObjectiveRFun$new(
  fun = objective_function,
  domain = domain,
  codomain = codomain,
  id = "optimize_grl_weights_cmaes"
)

instance = OptimInstanceSingleCrit$new(
  objective = objective,
  terminator = trm("evals", n_evals = 100L)
)

optimizer = opt("cmaes")
optimizer$optimize(instance)

# configure the ensemble with the optimized weights and evaluate on the outer split
grl_weights_optimized_cmaes = grl$clone(deep = TRUE)
grl_weights_optimized_cmaes$param_set$set_values(.values = list(classifavg.weights = instance$archive$best()$weights[[1L]]))
grl_weights_optimized_cmaes$id = paste0(grl_weights_optimized_cmaes$id, ".optimized_cmaes")

bg = benchmark_grid(task, list(grl, grl_weights_optimized_cmaes, dt, kknn, log_reg, featureless), outer_split)
b = benchmark(bg)
b$aggregate(msr("classif.auc"))

### GES implementation

# selected is an integer vector of length t where t is the current iteration within GES
# learner_num_ids is an integer vector from 1, ..., M where M is the number of models
# get_ges_weights should calculate the weights for each model based on the number of times is has been selected
get_ges_weights = function(selected, learner_num_ids) {
  ...
  weights
}

# learner_preds is a list of the resampled learner predictions, e.g., list(dt_pred, kknn_pred, log_reg_pred)
# iterations is an integer specifying the number of iterations GES should be performed
# measure is the measure used to score the predictions
ges = function(learner_preds, iterations = 100L, measure = msr("classif.auc")) {
  iteration = 0
  learner_num_ids = seq_along(learner_preds)
  performance = numeric(iterations)
  selected = integer(iterations)
  for (iteration in seq_len(iterations)) {
    cat("Iteration: ", iteration, "\n")
    performance_tmp = map_dbl(learner_num_ids, function(learner_num_id) {
      # for each model: add it to the ensemble and calculate the average validation performance of the k folds by performing model averaging
      # use get_ges_weights to calculate the weights for each model
      selected_tmp = selected
      selected_tmp[iteration] = learner_num_id
      weights = get_ges_weights(...)
      performances = map_dbl(seq_len(k), function(fold) {
        # similar steps as in the ObjectiveRFun we constructed to optimized the weights numerically with CMA-ES
        ...
      })
      mean(performances)
    })
    cat("Performance if learner added: ", round(performance_tmp, 3), "\n")
    if (!measure$minimize) {
      performance_tmp = - performance_tmp
    }
    select = ... # best model if added to the ensemble
    performance[iteration] = performance_tmp[select]
    selected[iteration] = select
  }
  best_iteration = which.min(performance)
  weights = get_ges_weights(selected[seq_len(best_iteration)], learner_num_ids = learner_num_ids)
  if (!measure$minimize) {
    performance = - performance
  }
  list(best_iteration = best_iteration, performance_best = performance[best_iteration], weights = weights, selected = selected, performance = performance)
}

ges_results = ges(list(dt_pred, kknn_pred, log_reg_pred))

# configure the ensemble with the optimized weights and evaluate on the outer split
grl_weights_optimized_ges = grl$clone(deep = TRUE)
grl_weights_optimized_ges$param_set$set_values(.values = list(classifavg.weights = ...))
grl_weights_optimized_ges$id = paste0(grl_weights_optimized_ges$id, ".optimized_cmaes")

bg = benchmark_grid(task, list(grl, grl_weights_optimized_cmaes, grl_weights_optimized_ges, dt, kknn, log_reg, featureless), outer_split)
b = benchmark(bg)
b$aggregate(msr("classif.auc"))
```
</details>

```{r, eval = !show.solution, echo = FALSE, results='asis'}
cat("<!--")
```
### Solution:

<details>
<summary>Click me:</summary>

```{r, eval = show.solution}
library(mlr3)
library(mlr3learners)
library(mlr3pipelines)
library(mlr3viz)
library(bbotk)
library(mlr3misc)

task = tsk("german_credit")

outer_split = rsmp("holdout")$instantiate(task)

set.seed(2409)
dt = lrn("classif.rpart")
dt$predict_type = "prob"
kknn = lrn("classif.kknn", k = 7L)
kknn$predict_type = "prob"
log_reg = lrn("classif.log_reg")
log_reg$predict_type = "prob"
featureless = lrn("classif.featureless")
featureless$predict_type = "prob"

gr = gunion(list(dt, kknn, log_reg)) %>>% po("classifavg")
grl = as_learner(gr)

bg = benchmark_grid(task, list(grl, dt, kknn, log_reg, featureless), outer_split)
b = benchmark(bg)
b$aggregate(msr("classif.auc"))

# optimize weights on the train/validation split from the outer_split
train_valid_task = task$clone(deep = TRUE)$filter(outer_split$train_set(1L))
k = 10L
resampling = rsmp("cv", folds = k)$instantiate(train_valid_task)

dt_pred = resample(train_valid_task, dt, resampling)
kknn_pred = resample(train_valid_task, kknn, resampling)
log_reg_pred = resample(train_valid_task, log_reg, resampling)

objective_function = function(xs) {
  weights = unlist(xs)
  weights = weights / sum(weights)  # sum to 1 normalization
  aucs = map_dbl(seq_len(k), function(fold) {
    dt_p = dt_pred$predictions()[[fold]]
    kknn_p = kknn_pred$predictions()[[fold]]
    log_reg_p = log_reg_pred$predictions()[[fold]]
    row_ids = dt_p$row_ids
    stopifnot(all(row_ids == kknn_p$row_ids) && all(row_ids == log_reg_p$row_ids))
    truth = dt_p$truth
    weighted_probs = Reduce("+", list(dt_p$prob * weights[1L], kknn_p$prob * weights[2L], log_reg_p$prob * weights[3L]))
    weighted_response = colnames(weighted_probs)[apply(weighted_probs, MARGIN = 1L, FUN = which.max)]
    weighted_p = PredictionClassif$new(row_ids = row_ids, truth = truth, response = weighted_response, prob = weighted_probs)
    weighted_p$score(msr("classif.auc"))
  })
  list(classif.auc = mean(aucs), weights = list(weights))
}

domain = ps(w_dt = p_dbl(0, 1), w_kknn = p_dbl(0, 1), w_log_reg = p_dbl(0, 1))
codomain = ps(classif.auc = p_dbl(0.5, 1, tags = "maximize"))

objective = ObjectiveRFun$new(
  fun = objective_function,
  domain = domain,
  codomain = codomain,
  id = "optimize_grl_weights_cmaes"
)

instance = OptimInstanceSingleCrit$new(
  objective = objective,
  terminator = trm("evals", n_evals = 100L)
)

optimizer = opt("cmaes")
optimizer$optimize(instance)

# configure the ensemble with the optimized weights and evaluate on the outer split
grl_weights_optimized_cmaes = grl$clone(deep = TRUE)
grl_weights_optimized_cmaes$param_set$set_values(.values = list(classifavg.weights = instance$archive$best()$weights[[1L]]))
grl_weights_optimized_cmaes$id = paste0(grl_weights_optimized_cmaes$id, ".optimized_cmaes")

bg = benchmark_grid(task, list(grl, grl_weights_optimized_cmaes, dt, kknn, log_reg, featureless), outer_split)
b = benchmark(bg)
b$aggregate(msr("classif.auc"))

get_ges_weights = function(selected, learner_num_ids) {
  selected = selected[selected != 0]
  n = length(selected)
  weights = map_dbl(learner_num_ids, function(learner_num_id) {
    sum(learner_num_id == selected) / n
  })
  weights
}

ges = function(learner_preds, iterations = 100L, measure = msr("classif.auc")) {
  iteration = 0
  learner_num_ids = seq_along(learner_preds)
  performance = numeric(iterations)
  selected = integer(iterations)
  for (iteration in seq_len(iterations)) {
    cat("Iteration: ", iteration, "\n")
    performance_tmp = map_dbl(learner_num_ids, function(learner_num_id) {
      selected_tmp = selected
      selected_tmp[iteration] = learner_num_id
      weights = get_ges_weights(selected_tmp, learner_num_ids = learner_num_ids)
      performances = map_dbl(seq_len(k), function(fold) {
        learner_ps = map(learner_preds, function(learner_pred) learner_pred$predictions()[[fold]])
        row_ids = learner_ps[[1L]]$row_ids
        stopifnot(all(map_lgl(learner_ps[-1L], function(learner_p) all(learner_p$row_ids == row_ids))))
        truth = learner_ps[[1L]]$truth
        weighted_probs = Reduce("+", mapply(FUN = function(prob, weight) prob * weight, map(learner_ps, function(learner_p) learner_p$prob), as.list(weights), SIMPLIFY = FALSE))
        stopifnot(all.equal(rowSums(weighted_probs),  rep(1, nrow(weighted_probs))))
        weighted_response = colnames(weighted_probs)[apply(weighted_probs, MARGIN = 1L, FUN = which.max)]
        weighted_p = PredictionClassif$new(row_ids = row_ids, truth = truth, response = weighted_response, prob = weighted_probs)
        weighted_p$score(measure)
      })
      mean(performances)
    })
    cat("Performance if learner added: ", round(performance_tmp, 3), "\n")
    if (!measure$minimize) {
      performance_tmp = - performance_tmp
    }
    select = which.min(performance_tmp)
    performance[iteration] = performance_tmp[select]
    selected[iteration] = select
  }
  best_iteration = which.min(performance)
  weights = get_ges_weights(selected[seq_len(best_iteration)], learner_num_ids = learner_num_ids)
  if (!measure$minimize) {
    performance = - performance
  }
  list(best_iteration = best_iteration, performance_best = performance[best_iteration], weights = weights, selected = selected, performance = performance)
}

ges_results = ges(list(dt_pred, kknn_pred, log_reg_pred))

# configure the ensemble with the optimized weights and evaluate on the outer split
grl_weights_optimized_ges = grl$clone(deep = TRUE)
grl_weights_optimized_ges$param_set$set_values(.values = list(classifavg.weights = ges_results$weights))
grl_weights_optimized_ges$id = paste0(grl_weights_optimized_ges$id, ".optimized_cmaes")

bg = benchmark_grid(task, list(grl, grl_weights_optimized_cmaes, grl_weights_optimized_ges, dt, kknn, log_reg, featureless), outer_split)
b = benchmark(bg)
b$aggregate(msr("classif.auc"))
# We already discussed that GES effectively optimizes weights on an integer scale
# CMA-ES in contrast can overfit to the validation data
# For some more background, see, e.g., https://arxiv.org/pdf/2307.00286

```
```{r, eval = !show.solution, echo = FALSE, results='asis'}
cat("-->")
```

## Exercise 2: Compare Weighted Model Averaging to Weighted Majority Voting (Hard Voting)

So far, we always performed weighted model averaging to build our ensemble.
More precisely, we required probability predictions of each learner and weighted these probabilities and summed these weighted probabilities over all learners.
To obtain a hard label response for each input, we took the class label that has the highest average weighted probability.

Now, we also want to take a look at (weighted) majority voting (also known as hard voting).
Note that the classifavg `PipeOp` can actually be used to perform both model averaging and majority voting.
Its behavior simply depends on the `$predict_type` of the learners.
If the learners return probability predictions, model averaging is performed.
If the learners return responses (hard labels), majority voting is performed.

Perform weighted majority voting (equal weights) of a decision tree, a k-NN (k = 7) and a logistic regression
and compare this to weighted model averaging (equal weights) of the same learners.

Use the outer test split from above as a resampling and evaluate models with respect to accuracy.

What happens if we now want to evaluate with respect to ROC AUC?
How can the weighted majority voting ensemble still be benchmarked with respect to ROC AUC?
(We would expect it to simple return a hard label response, right?)
Check the probability predictions of the ensemble.

<details>
  <summary>**Hint 1:**</summary>
  To understand how model averaging vs. majority voting is implemented, see https://github.com/mlr-org/mlr3pipelines/blob/HEAD/R/PipeOpClassifAvg.R and the helper code in
  https://github.com/mlr-org/mlr3pipelines/blob/20486ca70d975e1a5a03472a1d5eb843256aa05e/R/PipeOpEnsemble.R#L174 (`weighted_matrix_sum` and `weighted_factor_mean`).
 
</details>

<details>
  <summary>**Hint 2:**</summary>
 
```{r, eval = FALSE}
dt_resp = lrn(...)
kknn_resp = lrn(...)
log_reg_resp = lrn(...)

gr_resp = gunion(list(...)) %>>% po(...)
grl_resp = as_learner(gr_resp)
grl_resp$id = "Majority Voting"

dt_prob = lrn(...)
dt_prob$predict_type = "prob"
.
.
.

gr_prob = gunion(list(...)) %>>% po(...)
grl_prob = as_learner(gr_prob)
grl_prob$id = "Model Averaging"

bg = benchmark_grid(task, list(grl_resp, grl_prob), outer_split)
b = benchmark(bg)
b$aggregate(msr("classif.acc"))
b$aggregate(msr("classif.auc"))

.
.
.
```
</details>

```{r, eval = !show.solution, echo = FALSE, results='asis'}
cat("<!--")
```
### Solution:

<details>
<summary>Click me:</summary>

```{r, eval = show.solution}
dt_resp = lrn("classif.rpart")
kknn_resp = lrn("classif.kknn", k = 7L)
log_reg_resp = lrn("classif.log_reg")

gr_resp = gunion(list(dt_resp, kknn_resp, log_reg_resp)) %>>% po("classifavg")
grl_resp = as_learner(gr_resp)
grl_resp$id = "Majority Voting"

dt_prob = lrn("classif.rpart")
dt_prob$predict_type = "prob"
kknn_prob = lrn("classif.kknn", k = 7L)
kknn_prob$predict_type = "prob"
log_reg_prob = lrn("classif.log_reg")
log_reg_prob$predict_type = "prob"

gr_prob = gunion(list(dt_prob, kknn_prob, log_reg_prob)) %>>% po("classifavg")
grl_prob = as_learner(gr_prob)
grl_prob$id = "Model Averaging"

bg = benchmark_grid(task, list(grl_resp, grl_prob), outer_split)
b = benchmark(bg)
b$aggregate(msr("classif.acc"))
b$aggregate(msr("classif.auc"))  # interestingly, this works but we get a warning

# although we did not properly set the predict type of the ensemble, it still returns probability predictions
grl_resp$train(task)
grl_resp$predict(task)
# but the probabilities are just 0 or 1

# be careful when using a hard voting ensemble and its probabilities predictions

# also note that if you change
grl_resp$predict_type = "prob"
# the ensemble will just behave as a model averaging ensemble
bg = benchmark_grid(task, list(grl_resp, grl_prob), outer_split)
b = benchmark(bg)
b$aggregate(msr("classif.auc")) 
# because this also changed the predict type of each learner
grl$graph$pipeops$classif.rpart$predict_type
grl$graph$pipeops$classif.kknn$predict_type
grl$graph$pipeops$classif.log_reg$predict_type
```
```{r, eval = !show.solution, echo = FALSE, results='asis'}
cat("-->")
```

# Summary
We have taken a deeper dive into greedy ensemble selection and implemented it from scratch.
Moreover, we compared model averaging to majority voting.

