---
title: "Hypothesis Testing for Benchmarking"
output:
  prettydoc::html_pretty:
    theme: hpstr
    toc: yes
    toc_depth: 2
    css: ../_template/prettydoc-hpstr.css
    self_contained: yes
---

```{r, echo = FALSE}
show.solution = TRUE
```

```{r, eval = TRUE, echo = FALSE, include=FALSE}
sys.source("../setup/setup.R", envir = knitr::knit_global())
set.seed(123)
library(mlr3oml)
library(mlr3verse)
library(mlr3learners)
library(tidyverse)
library(ggplot2)
```

# Goals

Our goal for this exercise sheet is to use `mlr3` to benchmark models in multiple scenarios, using hypothesis tests as an additional diagnostic.

## Question 1: 2 Algorithms on 1 Data Set

#### (a) Train a random forest from the `ranger` package and a regression tree from the `rpart` package using `mlr3` with default hyperparameters on the German credit task ("german_credit").

```{r, eval = show.solution, echo = show.solution}

rpart.model = mlr_learners$get("classif.rpart")
rpart.model$predict_type = "prob"
library(mlr3verse)
library(tidyverse)
library(ggplot2)
credit.task = tsk("german_credit")
set.seed(20220801)

rpart.model$train(credit.task)

ranger.model = mlr_learners$get("classif.ranger")
ranger.model$predict_type = "prob"
ranger.model$train(credit.task)
```

#### (b) Create a matrix with the ground truth label, the predicted probabilities by rpart and ranger, and the predicted labels.

```{r, eval = show.solution, echo = show.solution}
rpart.predictions = rpart.model$predict_newdata(newdata = credit.task$data())$print()
ranger.predictions = ranger.model$predict_newdata(newdata = credit.task$data())$print()
conf.mat = rpart.predictions[ , c("row_ids", "truth", "response", "prob.good")]
conf.mat$ranger = ranger.predictions$response
conf.mat$ranger_prob = ranger.predictions$prob.good
colnames(conf.mat) = c("id", "truth", "rpart", "rpart_prob", "ranger", "ranger_prob")
conf.mat
```

#### (c) Add two new columns with the observation-wise loss value and compare the performance of both models using the Brier score.

```{r, eval = show.solution, echo = show.solution}
conf.mat = conf.mat %>%
  mutate(
    rpart_loss = (rpart_prob - ifelse(truth == "good", 1, 0))^2
  ) %>%
  mutate(
    ranger_loss = (ranger_prob - ifelse(truth == "good", 1, 0))^2
  ) %>%
  mutate(
    diff_loss = rpart_loss - ranger_loss
  )

brier_score_ranger = mean(conf.mat$ranger_loss)
brier_score_rpart = mean(conf.mat$rpart_loss)

```

#### (d) Use a two sample t-test to evaluate whether both samples of performance scores come from different populations.

Hint: Add another column with the difference between observation-wise loss values. Then run the t-test.

```{r, eval = show.solution, echo = show.solution}
conf.mat = conf.mat %>%
  mutate(
    diff_loss = rpart_loss - ranger_loss
  )

mean_diff_loss = conf.mat %>%
  ungroup() %>%
  summarize(diff_loss = mean(diff_loss))
mean_diff_loss

t_var = sqrt((1 / (nrow(conf.mat) - 1)) * sum((conf.mat$diff_loss - as.numeric(mean_diff_loss))^2))
t_var

t_statistic = sqrt(nrow(conf.mat)) * (mean_diff_loss / t_var)
t_statistic
# lower critical value, reject H0 if t-stat smaller
qt(0.025, df = nrow(conf.mat) - 1)
# OR
# upper critical value, reject H0 if t-stat larger
qt(0.025, df = nrow(conf.mat) - 1)

t_statistic > qt(0.025, df = nrow(conf.mat) - 1)
```

<!-- We reject the null. This means the difference in model performance is significant. -->

#### (d) Now run the McNemar test.

Hint: You will need the total number of observations that are classified correctly by rpart only and that are classified correctly by ranger only.


```{r, eval = show.solution, echo = show.solution}

conf.mat = conf.mat %>%
  rowwise() %>%
  mutate(
    rpart_correct = case_when(
      truth == "good" & rpart == "good" |
        truth == "bad" & rpart == "bad"
      ~ 1,
      .default = 0)) %>% 
  mutate(
    ranger_correct = case_when(
      truth == "good" & ranger == "good" |
        truth == "bad" & ranger == "bad"
      ~ 1,
      .default = 0)) %>%
  mutate(
    only_rpart_correct = case_when(
      rpart_correct == 1 & ranger_correct == 0
      ~ 1,
      .default = 0)) %>% 
  mutate(
    only_ranger_correct = case_when(
      rpart_correct == 0 & ranger_correct == 1
      ~ 1,
      .default = 0))


n_only_rpart_correct = sum(conf.mat$only_rpart_correct)
n_only_ranger_correct = sum(conf.mat$only_ranger_correct)

mcnemar_stat = (abs(n_only_rpart_correct - n_only_ranger_correct) - 1)^2 / (n_only_rpart_correct + n_only_ranger_correct)
mcnemar_stat

# critical value for alpha = 5%, reject H0 if mcnemar_stat larger
crit = qchisq(0.95, df = 1)

mcnemar_stat > crit
```

<!-- We reject the null. This means that also for the McNemar test, the difference in performance is statistically significant. -->

## Question 2: 2 Algorithms on Multiple Data Sets

Let us now confirm whether this result holds for other data sets as well. We will first scout OpenML (package `mlr3oml`) for suitable classification tasks. 

#### (a) Use the function `list_oml_tasks` to look for tasks with the following characteristics: 

- binary classification
- number of features between 5 and 10
- number of instances between 500 and 10000
- no missing values

```{r, eval = show.solution, echo = show.solution, results=FALSE}
library(mlr3oml)
otasks = list_oml_tasks(
  type = "classif",
  number_features = c(5, 10),
  number_instances = c(500, 10000),
  number_classes = 2,
  number_missing_values = 0
)
```


#### (b) Filter out tasks with the same data_id. Use only tasks with balanced data sets where the ratio between the majority and minority class is smaller than 1.2. Also, remove tasks with a data_id of 720. You should receive a total number of 20 tasks. 

```{r, eval = show.solution, echo = show.solution}
otasks = otasks %>%
  filter(data_id != 720) %>%
  filter(MajorityClassSize / MinorityClassSize < 1.2) %>%
  group_by(name) %>%
  filter(row_number(data_id) == 1) %>%
  group_by(data_id) %>%
  filter(row_number(data_id) == 1)

nrow(otasks)
```

#### (c) Benchmark rpart and ranger on all found tasks with mlr3. Use one-hot encoding and three-fold cross-validation.

Hint: `po("encode", method = "one-hot")`

```{r, eval = show.solution, echo = show.solution}

tasklist = lapply(otasks$task_id, FUN = function(x) otsk(x))
unlist(lapply(tasklist, FUN = function(x) x[["data_name"]]))

learners = lapply(c("classif.rpart", "classif.ranger"), lrn)
poe = po("encode", method = "one-hot")
learners = lapply(learners, FUN = function(x) {po("encode") %>>% x})

resamplings = rsmp("cv", folds = 3)
design = benchmark_grid(tasklist, learners, resamplings)
set.seed(123)
bmr = benchmark(design)
```

#### (d) Apply the `$aggregate()` function to the `mlr3` benchmark object and compare the ranks of both algorithms on all tasks.

```{r, eval = show.solution, echo = show.solution}
aggr = bmr$aggregate()
aggr = aggr %>%
  mutate(learner_id = replace(learner_id, learner_id == "encode.classif.ranger", "ranger")) %>%
  mutate(learner_id = replace(learner_id, learner_id == "encode.classif.rpart", "rpart"))

ranktable_wide_rpart_ranger = aggr %>%
  group_by(task_id) %>%
  mutate(rank_on_task = rank(classif.ce)) %>%
  mutate(ce_rank = rank_on_task) %>%
  select(c(task_id, learner_id, ce_rank)) %>%
  pivot_wider(names_from = learner_id, values_from = ce_rank)

ranktable_wide_rpart_ranger
```

#### (e) Run the Wilcoxon signed rank test using the ranks you computed:

```{r, eval = show.solution, echo = show.solution}
r_plus = sum(as.numeric(ranktable_wide_rpart_ranger$rpart == 1))
r_minus = sum(as.numeric(ranktable_wide_rpart_ranger$rpart == 2))
M = nrow(ranktable_wide_rpart_ranger)
wilcoxon = (min(r_plus, r_minus) - ((1 / 4) * M * (M + 1))) / sqrt((1 / 24) * M * (M + 1) * (2 * M + 1)) 
wilcoxon
```

## Question 3: Multiple Algorithms on Multiple Data Sets

#### (a) Let us now compare more algorithms on each task. Rerun the benchmark with the learners "classif.featureless", "classif.cv_glmnet", "classif.rpart", "classif.ranger", "classif.kknn", and "classif.svm". As before, use one-hot encodings.

```{r, eval = show.solution, echo = show.solution}

learners = lapply(c("classif.featureless",  "classif.cv_glmnet", "classif.rpart", "classif.ranger", "classif.kknn", "classif.svm"), lrn)
poe = po("encode", method = "one-hot")
learners = lapply(learners, FUN = function(x) {po("encode") %>>% x})

resamplings = rsmp("cv", folds = 3)
design = benchmark_grid(tasklist, learners, resamplings)

set.seed(123)
bmr = benchmark(design)

aggr = bmr$aggregate()
aggr = aggr %>%
  mutate(learner_id = replace(learner_id, learner_id == "encode.classif.featureless", "featureless")) %>%
  mutate(learner_id = replace(learner_id, learner_id == "encode.classif.ranger", "ranger")) %>%
  mutate(learner_id = replace(learner_id, learner_id == "encode.classif.kknn", "kknn")) %>%
  mutate(learner_id = replace(learner_id, learner_id == "encode.classif.rpart", "rpart")) %>%
  mutate(learner_id = replace(learner_id, learner_id == "encode.classif.svm", "svm")) %>%
  mutate(learner_id = replace(learner_id, learner_id == "encode.classif.cv_glmnet", "cv_glmnet"))
```

#### (b) Given multiple algorithms on multiple data sets, we have to use an omnibus test such as the Friedman test. Compute a rank table that tells you the rank of each algorithm for each task. Then, compute the average rank of each algorithm and proceed with the computation of the Friedman statistic.

```{r, eval = show.solution, echo = show.solution}

ranktable = aggr %>%
  group_by(task_id) %>%
  mutate(rank_on_task = rank(classif.ce)) %>%
  mutate(ce_rank = paste(round(classif.ce, 4), " (", rank_on_task, ")", sep =  ""))

averageranks = ranktable %>%
  group_by(learner_id) %>%
  summarize(average_rank_on_task = mean(rank_on_task))

averageranks

meanrank = (1 / nrow(ranktable)) * sum(ranktable$rank_on_task)
meanrank

sstotal = length(tasklist) *
  sum((averageranks$average_rank_on_task- meanrank)^2)
sstotal

sserror = (1 / (length(tasklist) * (length(learners) - 1))) * sum((ranktable$rank_on_task - meanrank)^2)
sserror

friedmanstat = sstotal / sserror
friedmanstat
```

#### (c) Run a sanity check with the `friedman.test` function implemented in the `stats` package.

```{r, eval = show.solution, echo = show.solution}
ce_table_wide = ranktable %>%
  select(c(task_id, learner_id, classif.ce)) %>%
  pivot_wider(names_from = learner_id, values_from = classif.ce) %>%
  ungroup() %>%
  select(-task_id) %>%
  as.matrix()
ce_table_wide

friedman.test(ce_table_wide)
```

#### (d) Use the function `frdAllPairsNemenyiTest` from the `PMCMRplus` package to run all pairwise Nemenyi tests.

<!-- The Friedman test indicates that at least one algorithm performs differently. We can run pairwise comparisons with post-hoc tests such as the Nemenyi or Bonferroni-Dunn test. -->

```{r, eval = show.solution, echo = show.solution}
library(PMCMRplus) 
rownames(ce_table_wide) = LETTERS[1:nrow(ce_table_wide)]
nemenyi_test = frdAllPairsNemenyiTest(y = ce_table_wide)
nemenyi_test$statistic
nemenyi_test$p.value
```

#### (e) Manually compute the critical difference for rpart and ranger.

Hint: Use the `qtukey` function.

```{r, eval = show.solution, echo = show.solution}
mean_diff_rpart_ranger = averageranks[averageranks$learner_id == "rpart", 2] - averageranks[averageranks$learner_id == "ranger", 2]
# # critical mean rank difference
crit_value_mean_rank_diff = (qtukey(p = 0.05, df = Inf, nmeans = 6, lower.tail = FALSE) / sqrt(2)) * (sqrt((6 * (6 + 1)) / (6 * 16)))

mean_diff_rpart_ranger > crit_value_mean_rank_diff
# reject H_0!
```

#### (f) Manually compare `rpart` and `ranger` with the Bonferroni-Dunn test. 

Hint: The probability of observing the test statistic under the null hypothesis is given by `pnorm(<...>, 0, 1, lower.tail = FALSE)`.

```{r, eval = show.solution, echo = show.solution}
# bonferroni dunn test
dunn_stat = (abs(mean_diff_rpart_ranger$average_rank_on_task) / sqrt((6 * 7) / 6 * 16))
# prob of observing dunn statistic under h0:
pnorm(dunn_stat, 0, 1, lower.tail = FALSE)
# do not reject H_0
```
<!-- Interestingly, both tests differ in this case. The Nemenyi test lets us reject the null, while the Bonferroni-Dunn test does not let us reject the null. -->


#### (g) Compute a criticial difference plot with `mlr3`.

```{r, eval = show.solution, echo = show.solution}
# sanity check with mlr3
library(mlr3benchmark)
benchmark_aggr = as_benchmark_aggr(bmr, measures = msr("classif.ce"))
benchmark_aggr$friedman_test()
cd_plot = autoplot(benchmark_aggr, type = "cd", meas = "ce", minimize = TRUE)
cd_plot
```
